### 2016-06-20 ansi-isolation.tex
Example added.

ANSI SQL 92 standard requires \iso{Read Committted} isolation to avoid
dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. \C{RC}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of RC for highly available
replicated stores. On relational databases, however, RC has also come
to be associated with the $\mathtt{MonotonicVis}$ guarantee.
Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ are
sufficient to reason about RC isolation on relational stores too. As
we demonstrate in \S\ref{sec:store-consistency}, the combination of
these guarantees with the SC property of relational stores
automatically leads to the monotonicity guarantee, which probably
explains why RC comes with $\mathtt{MonotonicVis}$ on such stores. On
weakly consistent stores however, $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ do not imply $\mathtt{MonotonicVis}$. A stronger
isolation level called \emph{Monotonic Atomic View} (\C{MAV} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend RC with monotonicity on such stores. \iso{Repeatable
Read} spec (\C{RR}) extends \C{MAV} with snapshot visibility guarantee
described before. \iso{Snapshot Isolation} spec (\C{SI}) extends
\C{RR} with one-way serializability guarantee w.r.t the transactions
that perform conflicting writes (i.e., writes to the same shared
variable). Finally, \iso{Serializable} isolation extends one-way
$\mathtt{SnapshotVis}$ to all transactions, including those that do
not conflict.

It is worth noting that, unlike some previous proposals
~\cite{gotsmanconcur15}, our specifications for SI and SER do not
necessarily impose a total order among transactions (conflicting or
otherwise). In reality, total order under SER (resp. SI) is guaranteed
only if the store executes all transactions under SER (resp. SI)
isolation. Our specifications admit this possibility, and derive total
order under the assumption of homogenity. However, databases almost
always allow isolation levels to be configured on a per-transaction
basis, allowing transactions at various isolation levels to coexist.
Specifications that assume homogenity are incorrect under this
setting\footnote{On the other hand, Fig.~\ref{fig:ansi-isolation}
presents slightly weaker versions of \C{SI} and \C{SER} specs in the
interest of clarity.}.

>>>>

ANSI SQL 92 standard requires \iso{Read Committted} isolation to avoid
dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. \C{RC}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of RC for highly available
replicated stores. On relational databases, however, RC has also come
to be associated with the $\mathtt{MonotonicVis}$ guarantee.
Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ are
sufficient to reason about RC isolation on relational stores too. As
we demonstrate in \S\ref{sec:store-consistency}, the combination of
these guarantees with the SC property of relational stores
automatically leads to the monotonicity guarantee, which probably
explains why RC comes with $\mathtt{MonotonicVis}$ on such stores. On
weakly consistent stores however, $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ do not imply $\mathtt{MonotonicVis}$. A stronger
isolation level called \iso{Monotonic Atomic View} (\C{MAV} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend RC with monotonicity on such stores. \iso{Repeatable
Read} spec (\C{RR}) extends \C{MAV} with snapshot visibility guarantee
described before. \iso{Snapshot Isolation} spec (\C{SI}) extends
\C{RR} with one-way serializability guarantee w.r.t the transactions
that perform conflicting writes (i.e., writes to the same shared
variable). Finally, \iso{Serializable} isolation extends one-way
$\mathtt{SnapshotVis}$ to all transactions, including those that do
not conflict\footnote{Fig.~\ref{fig:ansi-isolation} presents slightly
weaker versions of \C{SI} and \C{SER} specs in the interest of
clarity.}.

It is worth noting that, unlike some previous proposals
~\cite{gotsmanconcur15}, our specifications for SI and SER do not
necessarily impose a total order among transactions (conflicting or
otherwise). In reality, total order under SER (resp. SI) is guaranteed
only if the store executes all transactions under SER (resp. SI)
isolation. Our specifications admit this possibility, and derive total
order under the assumption of homogenity. However, databases almost
always allow isolation levels to be configured on a per-transaction
basis, allowing transactions at various isolation levels to coexist.
Specifications that assume homogenity are incorrect under this
setting.

Having specified isolation levels as trace well-formedness
constraints, we can now construct trace invariants ($\I$) for \txnimp
programs by composing isolation specifications for various
transactions. For instance, the following trace invariant enforces SI
for both transactions of the program in Fig.~\ref{fig:motiv-eg-1},
allowing it to satisfy its postcondition:
\begin{smathpar}
\I \;=\; \lambda\E.~ \underE{\C{SI(Wd1)}} \conj \underE{\C{SI(Wd2)}}
\end{smathpar}
In contrast, the following trace invariant enforces RC for one and SI
for another, leading to a possible violation of the postcondition:
\begin{smathpar}
\I \;=\; \lambda\E.~ \underE{\C{RC(Wd1)}} \conj \underE{\C{SI(Wd2)}}
\end{smathpar}
